---
title: "Time Series Analysis - Assigment 2"
author: "Loan Huynh - s3655461"
output:
  word_document: default
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

## Introduction

This report tries to find the suitable model which used to predict Egg deposition in next 5 years. The dataset is Egg depositions of age-3 Lake Huron Bloasters between 1981 and 1996 in **BloaterLH** dataset of **FSAdata** package

```{r warning=FALSE, message=FALSE}
# Loadpackage
library(FSAdata)
library(TSA)
library(fUnitRoots) # package used to test Non-stationary series
library(tseries)
library(lmtest)
library(forecast) # package used for forecasting
```
## Data Description

```{r}
# Load dataset
data(BloaterLH)

# Only get Eggs column data in BloaterLH 
egg <- BloaterLH$eggs

# Create Time Series objects for Egg data
egg_ts <- ts(egg, start = 1981, end = 1996)

# Plot Egg Deposition Series data
plot(egg_ts, type = 'o', xlab= 'Year',ylab="Egg Deposition", panel.first=grid(), main = "Time Series of Egg Deposition")
```

This series has an upward trend and changes of variation, but no repeating pattern in seasonality. There are some suceeding points but no fluctuation. It looks like the autoregressive </br>

Using Scatter Plot to see the relationship between Egg Deposition of previous years

```{r}
# Scatter Plot
plot(y= egg_ts , x = zlag(egg_ts ), ylab = 'Egg Deposition Series', xlab =  'Previous Year',
     panel.first=grid(), main = "Scatter Plot of Egg Deposition in Previous Years")
abline(coef = c(0,1), col= "red")
```
It is obviously to see that there is a strong correlation between Egg Depositions of previous Years in Scatter Plot. The following code is to calculate the value of correlation

```{r}
y <- egg_ts # Load data into y

x <- zlag(egg_ts) # Generate first lag of the abundance series

index <- 2:length(x)# Create index to get rid of the first NA value in x

cor(y[index], x[index])# Calculate the correlation between numerical values in x and y
```

The correlation is 0.7445657. This is a strong positive correlation </br>

Apply ADF Test to check whether series in stationary or non-stationary

```{r}
ar(diff(egg_ts))
```
Order of this test is 0

```{r}
adfTest(egg_ts, lags = 0)
```

The p_value is 0.452  larger than 5% level of significant. Therefore we cannot reject null hypothesis about non-stationary. It implies that this series is non-stationary. It is clearly shown in ACF and PACF plots as below:

```{r warning=FALSE, message=FALSE}
par(mfrow=c(1,2))
acf(egg_ts)
pacf(egg_ts)
```


The ACF plot has pattern which has a large spike at lag 1 that decrease after a few lags </br>

## DATA MODELING
#### DATA TRANSFORMATION AND DIFFERENCE

Apply Log transformation for this series

```{r}
# Using Log transform
log.egg <- log(egg_ts)
plot(log.egg , type = 'o', xlab="Year", ylab= "Egg Deposition", main =" Time Series of Transformed Egg Deposition", panel.first=grid())
```
This series still has a trend. </br>

Applying the first difference for transformed data

```{r}
diff.log.egg = diff(log.egg)
plot(diff.log.egg ,type='o',ylab='Egg Deposition', xlab= "Year", main = "Time Series - 1st Difference of Egg Deposition",
     panel.first=grid())
```

Applying ADF unit-root test to test the existence of non-stationary with this series

```{r}
ar(diff(diff.log.egg))

```

Order for this test is 4

```{r}
adfTest(diff.log.egg, lags = 4)
```
  
The p-value is 0.2049 larger than 5% level of significance. Therefore, this series is still non-stationary. Hence, we apply the second difference 

```{r}
diff2.log.egg <- diff(diff.log.egg, diference = 2)
plot(diff2.log.egg ,type='o',ylab='Egg Deposition', xlab= "Year", main = "Time Series - 2nd Difference of Egg Deposition",
     panel.first=grid())
```

Applying ADF unit-root test for second difference

```{r}
ar(diff(diff2.log.egg))
```

Order for this test is 1

```{r}
adfTest(diff2.log.egg, lags = 1)
```
p-value = 0.01 

The p-value equals 0.01 less than 5% level of signigicance. Therefore, we can reject null hypothesis for non-stationary. It means that this series is stationary. </br>

Displaying ACF and PACF plots

```{r}
par(mfrow=c(1,2))
acf(diff2.log.egg)
pacf(diff2.log.egg)
```
There is no pattern in ACF and PACF plots. </br>

__Using EACF__
```{r}
# Applying EACF
eacf(diff2.log.egg, ar.max = 3, ma.max = 3)
```
From the above matrix, the possible models are ARIMA(0,2,1), ARIMA(1,2,0), ARIMA(1,2,1) </br>

__Using BIC table__

```{r message=FALSE, warning=FALSE}
# Using BIC table
res = armasubsets(y=diff2.log.egg,nar=3,nma=3,y.name='test',ar.method='ols')
plot(res)
```

From BIC table, possible models include ARIMA(2,2,1) , ARIMA(2,2,3) and ARIMA(1,2,1)

Hence the set of possible models will be {ARIMA(2,2,1), ARIMA(1,2,1), ARIMA(2,2,3), ARIMA(0,2,1), ARIMA(1,2,0), ARIMA(1,2,1)}

##### PARAMETER ESTIMATION
In order to estimate parameters of possible models, I use Maximum Likelihood method </br>

__ARIMA(0,2,1)__

```{r}
# Maximum likelihood estimates of the coefficients for ARIMA(0,2,1)
model.021 <- arima(egg_ts,order=c(0,2,1),method='ML')
model.021
```
```{r}
# Maximum likelihood estimates of the coefficients with significance tests for ARIMA(0,2,1) model
coeftest(model.021)
```
The p-value is less than 0.05. Therefore the estimated coefficients of ARIMA(0,2,1) model is significant </br>

__ARIMA(2,2,1)__

```{r}
# Maximum likelihood estimates of the coefficients for ARIMA(2,2,1)
model.221 <- arima(egg_ts, order = c(2,2,1), method = 'ML')
model.221
```

```{r}
# Maximum likelihood estimates of the coefficients with significance tests for ARIMA(2,2,1) model
coeftest(model.221)
```

The estimated ar1 and ar2 coefficients are larger than 0.05. This means that they are are not significant

__ARIMA(1,2,1)__

```{r}
# Maximum likelihood estimates of the coefficients for ARIMA(1,2,1)
model.121 <- arima(egg_ts, order = c(1,2,1), method = 'ML')
model.121
```
```{r}
# Maximum likelihood estimates of the coefficients with significance tests for ARIMA(1,2,1) model
coeftest(model.121)
```

AR1 is insignificant in model ARIMA(1,2,1)

__ARIMA(2,2,3)__

```{r}
model.223 <- arima(egg_ts, order = c(2,2,3), method = 'ML')
coeftest(model.223)
```
AR1, AR2, MA1 and MA2 are insignificant

__ARIMA(1,2,0)__

```{r}
model.120 <- arima(egg_ts, order = c(1,2,0), method = 'ML')
coeftest(model.120)
```

AR1 is not significant as well in model ARIMA(1,2,0)


##### APPLYING AIC and BIC to decide the best model within the subset of possible models </br>

__ AIC__

```{r}
AIC(model.021)
```

```{r}
AIC(model.120)
```

```{r}
AIC(model.121)
```

```{r}
AIC(model.221)
```

```{r}
AIC(model.223)
```
__BIC__

```{r}
AIC(model.021, k = log(28))
```

```{r}
AIC(model.121, k = log(28))
```

```{r}
AIC(model.221, k = log(28))
```
```{r}
AIC(model.223, k = log(28))
```

The smallest AIC is model ARIMA(0,2,1) with  22.74602. According to BIC, model ARIMA(0,2,1) with value 25.41043 is the best. Therefore ARIMA(0,2,1) is the best model.

## MODEL DIAGNOTICS

__Normality Testing__

```{r}

plot(rstandard(model.021), type= 'o', ylab="Standardized Residuals", main = "Time Series Plot of standardised residuals for Egg Deposition Series")
abline(h=0)
```

```{r}
# Using QQplot
qqnorm(residuals(model.021))
qqline(residuals(model.021), col = 2)
```

Most of values lie on the red line. However, there are still some outliers at the top right and left bottom

```{r}
# Using Shapiro test
shapiro.test(diff2.log.egg)
```
The p-value is 0.531 > 0.05. Therefore, we cannot reject H0 hypothesis which is normality error assumption. This implies that normality error assumption is not violated </br>

__AutoCorrelation of residuals testing__

```{r}
acf(residuals(model.021))
```

```{r}
pacf(residuals(model.021))
```

From ACF and PACF plots of residuals, we can conclude that the residuals constitutes a white noise series as there is no highly significant correlation.

__LJUNG - BOX TEST__

```{r}
Box.test(residuals(model.021), lag = length(model.021$residuals) -1, type = "Box-Pierce", fitdf = 0)
```
The p-value in LjungBox test is 0.9892 > 0.05. Therefore, we cannot reject null hypothesis that the error terms are uncorrelated. </br>

In conclusion, there is no problem in the residuals of ARIMA(0,2,1) model.

## FORECASTING
```{r}
# Create a matrix of the covariance for prediction
xreg=data.frame (constant=seq(egg_ts))
n = length(egg_ts)
n.ahead = 5
newxreq = data.frame(constant = (n+1): (n+n.ahead))
# Predict Egg Deposition for next 5 years
predict(model.021, n.ahead = 5, newxreg = NULL, se.fit = TRUE)
```
The predict values of Egg depositions in next five years are {1.089693 ; 1.155287; 1.220880; 1.286473 ; 1.352067}
```{r}
#Plot forecasts over time series
fit = Arima(egg_ts,c(0,2,1))
plot(forecast(fit,h=10))
```


